### 2022-09-13 chapters/literature/draft_citations.tex
update :soa

as we see in % two diagrams one for research the other one for power usage 
The CPU is the part responsible of the most energy consumption of softwares. hence the finer we go to measure this energy consumption the better it is for our work.
Fortunately, Intel and later AMD proposed a tool that estimate the power consumption of different part of the CPI based on counter performances.
RAPL (\emph{RUnning Average Power Limit}) \cite{hackenberg2013power,hackenberg2015energy} is a set of registers that was introduced by Intel in their CPU since Sandy bridge generation, and later it was followed by AMD, since Family 17h Zen.

@@

jRAPL is another example of an energy measurement tool estimating tool that has been utilized in a variety of publications\cite{guimaraes2016some,liu2015data}.
This software enables for the energy usage of Java programs, functions, or even a block of code lines to be profiled and measured.
The measurements are heavily reliant on the data supplied by RAPL.
As a result, the global energy consumption collected by RAPL between two timestamps (the start and finish of the code to measure) is used to calculate the energy consumption of the Java code.
Tests using jRAPL should be done on a well-configured machine to minimize the impact of the operating system and user operations on the overall energy consumption of jRAPL.

Another process-level energy usage measuring tool is Jolinar~\cite{islam2016measuring,noureddine2016jolinar}.
The tool does not need a calibration phase and relies on pre-established power models based on hardware metrics (TDP, disk I/O rate, and so on).
These settings must be identified and supplied by the user for his machine.
Jolinar is only capable of measuring the energy consumption of one application at a time.
At the end of the execution, the tool provides the CPU, DRAM, and disk energy usage of the main process.

Jalen~\cite{noureddine2015monitoring} is another tool that profiles and monitors the energy usage of a Java program.
Unlike jRAPL, Jalen is able to cover the scope down to the function level.
It gathers data using code instrumentation and statistical sampling at a predetermined pace.
Because of the overhead that code instrumentation may incur, the authors recommend utilizing the second option.
Every 10 ms, Jalen records the JVM's stack trace together with the CPU time of threads and computes statistics about method calls.These statistics are then utilized to calculate each method's energy usage.

@@

%%****** mitigating the energy variation ******
low-cost scalable variation-aware algorithm \cite{inadomi2015analyzing}

reducing the energy variation by disabling turbo boost \cite{acun2016variation}

reducing the energy consumption in parallel systems \cite{chasapis2016runtime}

\subsection{Energy}
energy comparaison of programming languages in the game benchmark by debian \cite{pereira2017energy}

toward green ranking \cite{couto2017towards}

java vs kotlin in web performances \cite{bujnowski2020java}

rosetta code \cite{nanz2015comparative} \cite{mirowski2020rosetta}

network energy comparaison \cite{balasubramanian2009energy}

Aequitas \cite{ribic2016aequitas}

vm placement \cite{mishra2018energy}

\citeauthor{mishra2018energy}

static cost of Vms \cite{kurpicz2016much}

@@

%%%%%%%%%%%%% energy mitigation
\section{Energy Variation}\label{sec:energy_variation}
In theory, using identical CPU, same memory configuration, similar storage and networking capabilities, should increase the accuracy of physical measurements.
Unfortunately, this is not possible when it comes to measuring the energy consumption of a system.
Applying the benchmarking guidelines and repeating the same experiment with in the same configuration are not sufficient to reproduce the the same energy measurements, not only between identical machines, but even within the same machine.
This difference---also called \emph{energy variation} (EV)---has become a serious threat to the accuracy of experimental evaluations.

>>>>

As we see in % two diagrams one for research the other one for power usage 
The CPU is the part responsible for the most energy consumption in software. Hence, the finer we go to measure this energy consumption, the better it is for our work.
Fortunately, Intel and later AMD proposed a tool that estimates the power consumption of different parts of the CPI based on counter performances.
RAPL (\emph{RUnning Average Power Limit}) \cite{hackenberg2013power,hackenberg2015energy} is a set of registers that was introduced by Intel in their CPU since Sandy bridge generation, and later it was followed by AMD, since Family 17h Zen.

@@

JRAPL is another example of an energy measurement tool estimating tool that has been utilized in a variety of publications\cite{guimaraes2016some,liu2015data}.
This software enables for the energy usage of Java programs, functions, or even a block of code lines to be profiled and measured.
The measurements are heavily reliant on the data supplied by RAPL.
As a result, the global energy consumption collected by RAPL between two timestamps (the start and finish of the code to measure) is used to calculate the energy consumption of the Java code.
Tests using jRAPL should be done on a well-configured machine to minimize the impact of the operating system and user operations on the overall energy consumption of jRAPL.

Another process-level energy usage measuring tool is Jolinar~\cite{islam2016measuring,noureddine2016jolinar}.
The tool does not need a calibration phase and relies on pre-established power models based on hardware metrics (TDP, disk I/O rate, and so on).
These settings must be identified and supplied by the user for his machine.
Jolinar is only capable of measuring the energy consumption of one application at a time.
At the end of the execution, the tool provides the CPU, DRAM, and disk energy usage of the main process.
Jalen~\cite{noureddine2015monitoring} is another tool that profiles and monitors the energy usage of a Java program.
Unlike jRAPL, Jalen is able to cover the scope down to the function level.It gathers data using code instrumentation and statistical sampling at a predetermined pace.
Because of the overhead that code instrumentation may incur, the authors recommend utilizing the second option.
Every 10 ms, Jalen records the JVM's stack trace together with the CPU time of threads and computes statistics about method calls.These statistics are then utilized to calculate each method's energy usage.

@@

\section{Energy Variation}\label{sec:energy_variation}
In theory, using identical CPU, same memory configuration, similar storage and networking capabilities, should increase the accuracy of physical measurements.
Unfortunately, this is not possible when it comes to measuring the energy consumption of a system.
Applying the benchmarking guidelines and repeating the same experiment with in the same configuration are not sufficient to reproduce the the same energy measurements, not only between identical machines, but even within the same machine.
This difference---also called \emph{energy variation} (EV)---has become a serious threat to the accuracy of experimental evaluations.

\section{Energy optimisation}

We will now turn our focus to the energy optimisation challenge after considering the various ways for measuring energy consumption in computers and understanding the energy variation problem.
Over the previous decade, there has been a considerable increase of interest in this field, with many papers proposing different approaches to reduce the energy consumption of software applications. This section will passthrough the main contributions in this field, and will focus on the two following parts.

\subsection{Energy optimization in the conception phase}
The first part of this section will focus on energy optimisation in the conception phase, where the goal is to make the final product use less energy by choosing the best set of programming languages, tools, libraries, etc. It also includes all the work and optimizations that developers can do to the source code to make the software use less energy when it is running.

We start with the work of \citeauthor{pereira2017energy} where they  did an energy consumption comparison analysis of the most popular programming languages.
Using the pareto optimum, the paper  provides recommendations on how to combine some of these languages to enhance code quality while taking into account execution time, memory utilization, and energy consumption using the pareto optimum. Some of the study's findings demonstrate that interpreted languages, such as Python, have lower energy efficiency than compiled languages, such as C or Rust. The research also offers language combinations that developers might use together to improve energy economy, execution speed, and memory utilization.
This paper's findings were based on the game benchmark, which is the most famous set of benchmarks that compares several programming languages.

In this paper, we define a ranking of energy efficiency in programming languages. We considered a set of computing problems implemented in ten well-known programming languages and monitored the energy consumed when executing each language. Our preliminary results show that although the fastest languages tend to be the lowest energy-consuming ones, there are other interesting cases where slower languages are more energy-efficient than faster ones.

\citeauthor{couto2017towards} investigated the influence of programming language choice on the energy consumption of software during execution.In their paper~\cite{couto2017towards}, they examined a set of computing problems written in ten well-known programming languages while observing the energy required when running each language. They also found that there are interesting situations in which slower languages use less energy than faster ones, even though fast languages are usually the ones that use the least energy.
Finally, they produced an energy efficiency rating of programming languages.
The paper~\cite{bujnowski2020java} compared the energy, performance and database response time of web applications written in Java versus those written in Kotlin. They discovered that there is no statistically significant difference in CPU load between individual measurements( less than 2\%) 2. Howerver, Kotlin implementation has never earned the best results in any collection of measurements.

Other server works that have studied the impact of the website technology on the energy conusmptuon,\cite{philippot_characterization_2014},\cite{manotas2013investigating}. In their work~\cite{philippot_characterization_2014},\citeauthor{philippot_characterization_2014} measured the computer resources used during the loading of a website in a browser, such as memory utilization and energy consumption, for over 500 websites and proposed some best practices for developers.

% rosetta code \cite{nanz2015comparative} \cite{mirowski2020rosetta}

% network energy comparaison \cite{balasubramanian2009energy}

% Aequitas \cite{ribic2016aequitas}

% vm placement \cite{mishra2018energy}

% \citeauthor{mishra2018energy}

% static cost of Vms \cite{kurpicz2016much}

@@

%%%%%%%%%%%%% energy mitigation
%%****** mitigating the energy variation ******
% low-cost scalable variation-aware algorithm \cite{inadomi2015analyzing}

% reducing the energy variation by disabling turbo boost \cite{acun2016variation}

% reducing the energy consumption in parallel systems \cite{chasapis2016runtime}

### 2022-09-13 chapters/optimization/python-dir/python.tex
reformat text
update :soa

\subsection{Python Limits}
To reduce the energy consumption of Python, we started by targeting the main usage of this programming language, which is revealed to be data science and web development.
\Cref{fig:usecase} illustrates a study published by the JetBrain company on Python developers\footnote{\url{https://www.jetbrains.com/lp/python-developers-survey-2020}}.
57\% of the respondents reported that they use Python for data science, 51\% said they are using it for web development, and around 40\% are using it for system administration.
\note{The options in this survey were not mutually exclusive. As a result, the total of the percentages is greater than $100\%$.}

@@

In this chapter, we investigate the energy footprint of Python in its most popular domains of adoption.
We first explore the data and control structures, aiming to reveal some fundamental guidelines, as \citeauthor{s} did in~\cite{hasan_energy_2016-1}.
Then, we measure the energy consumption of several Python implementations to propose a non-intrusive technique to improve energy efficiency.
Therefore, this chapter will address the following research questions:
\begin{compactenum}[\indent\bf RQ\,1:]
    \item \emph{What is the energy footprint of Python when used in data science?}
    \item \emph{Are the Python guidelines energy efficient by construction?}
    \item \emph{Can we reduce the energy consumption of Python programs without altering the source code?}
\end{compactenum}

@@

\subsubsection{Key Performance Metrics}
We used Pytorch 1.10.0 to train those models and pyjoules to measure the energy consumption of the GPU and CPU.
\begin{itemize}
    \item accuracy : in \%
    \item execution time : in seconds for both the duration of each epoch and the total duration to a chieve a certain accuracy
    \item total energy consumption : for the CPU and the  GPU
    \item
\end{itemize}

>>>>

\subsection{Python Limits}
To reduce the energy consumption of Python, we started by targeting the main usage of this programming language, which is revealed to be data science and web development.
\Cref{fig:usecase} illustrates a study published by the JetBrain company on Python developers\footnote{\url{https://www.jetbrains.com/lp/python-developers-survey-2020}}.
57\% of the respondents reported that they use Python for data science, 51\% said they are using it for web development, and around 40\% are using it for system administration.
\paragraph{Remark}
The options in this survey were not mutually exclusive. As a result, the total of the percentages is greater than $100\%$.

@@

In this chapter, we investigate the energy footprint of Python in its most popular domains of adoption.
We first explore the data and control structures, aiming to reveal some fundamental guidelines, as \citeauthor{hasan_energy_2016-1} did in~\cite{hasan_energy_2016-1}.
Then, we measure the energy consumption of several Python implementations to propose a non-intrusive technique to improve energy efficiency.
Therefore, this chapter will address the following research questions:
\begin{compactenum}[\indent\bf RQ\,1:]
    \item \emph{What is the energy footprint of Python when used in data science?}
    \item \emph{Are the Python guidelines energy efficient by construction?}
    \item \emph{Can we reduce the energy consumption of Python programs without altering the source code?}
\end{compactenum}

@@

\subsubsection{Key Performance Metrics}
We used Pytorch 1.10.0 to train those models and pyjoules to measure the energy consumption of the GPU and CPU.
\begin{itemize}
    \item accuracy : in \%
    \item execution time : in seconds for both the duration of each epoch and the total duration to a chieve a certain accuracy
    \item total energy consumption : for the CPU and the  GPU
\end{itemize}
